A Lightweight, Swappable Voxel Ray-Tracer

### 2025-04-18 paper/paper.tex
updated paper

[Elaborate on voxel generation]

@@

In this paper, we aimed to compare two different storage solutions for a CPU voxel ray-tracer implemented in Rust. By comparing the average time it took to render randomly-generated scenes of varying sizes and at different resolutions, we found that a naive implementation (storing every voxel in an array, traversed with Amanatides and Woo’s “Fast Voxel Traversal” algorithm) was faster than a sparse storage solution (storing every voxel in an octree structure, traversed with a recursive algorithm) at smaller scene sizes, but scaled poorly and was significantly slower for the largest scenes that we tested. The sparse storage solution was faster for large scenes because its recursive structure eliminates some unnecessary traversal over empty space.

Some limitations and future areas of research for our study include but are not limited to the following: we only tested the ray tracer with relatively small scenes (<250 voxel extents), as it would have taken an impractical amount of time to do so with our computing resources. We also did not study the effects of using different visual textures.

\begin{thebibliography}{00}
\bibitem{williams} Williams, Amy \& Barrus, Steve \& Morley, R. \& Shirley, Peter. (2005). An Efficient and Robust Ray-Box Intersection Algorithm. J. Graphics Tools. 10. 49-54. 10.1145/1198555.1198748. 
\bibitem{amanatides} Amanatides, John \& Woo, Andrew. (1987). A Fast Voxel Traversal Algorithm for Ray Tracing. Proceedings of EuroGraphics. 87.
\end{thebibliography}

>>>>

Voxels are generated by the VoxelGenerator module. This module procedurally generates voxel data using Perlin noise. Perlin noise is a gradient-based noise function developed by Ken Perlin used in procedural generation due to its natural, smooth variation \cite{perlin} as seen in Figure ~\ref{perlin-noise}. The algorithm outputs pseudo-random values that transition smoothly across space, avoiding the harsh discontinuities of pure random noise.

\begin{figure}[htbp]
\centerline{\includegraphics[width=\linewidth]{perlin-noise}}
\caption{Example output of 2-D Perlin noise}
\label{perlin-noise}
\end{figure}

The module defines key structures and functions to determine voxel color and existence based on height-derived terrain. The module consists of the following structures and functions:

\cprotect\subsubsection{\verb|Voxel| struct}
The \verb|Voxel| struct encapsulates data for a single voxel. It contains one field:\verb|color| of type \verb|U8Vec3|: A 3-component vector representing RGB color values.

\cprotect\subsubsection{\verb|VoxelGenerator| struct}
The \verb|VoxelGenerator| struct is responsible for generating voxel data using Perlin noise. It contains \verb|perlin|: An instance of the \verb|Perlin| struct, which contains the perlin noise function seeded either randomly or explicitly.

The \verb|VoxelGenerator| can be initialized with either a random seed or a specific seed (mainly used for testing purposes)

Several constants are defined to control terrain parameters:
\begin{itemize}
\item \verb|HEIGHT|: Maximum height for voxel terrain.
\item \verb|ROUGHNESS|: Roughness factor for noise.
\item \verb|SCALE|: A scaling factor combining roughness and height to normalize noise frequency.
\end{itemize}

\cprotect\subsubsection{\verb|VoxelGenerator lookup()| function}
The lookup function computes Perlin noise at a given \verb|x,z| position by mapping the noise to a terrain height by scaling the values by the \verb|SCALE| constant. 
Then, a noise value is retrieved from the \verb|perlin.get()| function using the scaled coordinates.
The output of the noise function is in the range $[-1.0, 1.0]$. This value is normalized to the range $[0, \texttt{HEIGHT}]$ to compute the terrain height. It then compares the \verb|y| coordinate of the input position to the terrain height, and returns \verb|Some<Voxel>| if the position is at or below the terrain; otherwise returns \verb|None|.

\cprotect\subsubsection{\verb|VoxelGenerator  height_to_color()| function}
The function maps a height value to one of four constant RGB colors depending on the normalized terrain height, simulating terrain types like water, grass, mountains, and snow.

@@

[Draw some conclusion that affirms that we did what we set out to do]

\begin{thebibliography}{00}
\bibitem{williams} Williams, Amy \& Barrus, Steve \& Morley, R. \& Shirley, Peter. (2005). An Efficient and Robust Ray-Box Intersection Algorithm. J. Graphics Tools. 10. 49-54. 10.1145/1198555.1198748.
\bibitem{perlin} Perlin, Ken (1985). An image synthesizer. ACM SIGGRAPH Computer Graphics. 19 97–8930: 287–296. 10.1145/325165.325247
\bibitem{amanatides} Amanatides, John \& Woo, Andrew. (1987). A Fast Voxel Traversal Algorithm for Ray Tracing. Proceedings of EuroGraphics. 87.
\end{thebibliography}
